---
title: "Dataviz contest draft"
author: "C. Seth Lester, ASA"
date: "12/19/2020"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

# Data Exploration

Data visualization is a useful storytelling tool, but also plays an important role for the analyst in forming a comprehensive understanding of the data. In this section we will carry out the following tasks:

1. **Download the data** directly from the [SOA contest page](https://www.soa.org/research/opportunities/2020-data-visualization-contest/) to ensure that the most recent and updated data file is being used. Updates to field structure, factor delineation, or file type also will potentially cause program errors, which will serve to alert me to potential changes to the source data that may have occurred outside of my knowledge.
2. **Deflate the data** directly from the downloaded zipfile.
3. **Read and parse the data** from the relevant uncompressed CSV file.

## Download, uncompress, read and parse the SOA dataset

First we will load the CSV files listed on the SOA website. The complete end-to-end process to download, deflate, and parse the data from the CSV files offered by the SOA is detailed in the accompanying RMD file (code and output is suppressed to make for a prettier overall appearance of the final knitted document).

```{r load-data, include=FALSE}

source("./contest_loader.R")

```

## Data quality analysis

We owe the dataset an ASOP 23 once-over to determine:
1. Inconsistent values (e.g., in the sales data, date of the quote precedes app date)
2. Missing or problematic values (e.g. outliers)

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r eruptions, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r tabsets, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.



